---
title: "Monthly births QA"
output: html_notebook
---


Check that we have the expected year, gss code and gss name combinations for the monthly and aggregated monthly births

```{r setup}
library(tidyr)
library(dplyr)
source("R/functions/aggregate_to_combined_las.R")
```


```{r}
births_monthly_lad <- readRDS("data/processed/births_lad_monthly.rds")

geogs <- readRDS("lookups/lookup_lsoa_lad.rds") %>%
  select(gss_code, gss_name) %>%
  unique() %>%
  filter(grepl("^E", gss_code)) %>%
  filter(!gss_code %in% c("E09000012", "E09000001", "E06000052", "E06000053")) %>%
  bind_rows(data.frame(gss_code = c("E09000012, E09000001", "E06000052, E06000053"), 
                       gss_name = c("Hackney and City of London", "Cornwall and Isles of Scilly")))

sex <- "persons"
start_date <- min(births_monthly_lad$month_ending_date)
print(start_date)
end_date <- max(births_monthly_lad$month_ending_date)
print(end_date)

expected_months <- seq(start_date + 1, to = end_date + 1, by="month") - 1

expected_rows_df <- expand_grid(sex, month_ending_date = expected_months, nesting(gss_code = geogs$gss_code, gss_name = geogs$gss_name)) %>%
  mutate(type = "expected")

matched_df <- full_join(births_monthly_lad, expected_rows_df, by = join_by(gss_code, gss_name, month_ending_date, sex))

unmatched <- matched_df %>%
  filter(is.na(value) | is.na(type))

if (nrow(unmatched) != 0) stop("there are unmatched rows")
```



Check the monthly births data by comparing yearly aggregations to the annual births data

```{r read_data}
# Annual births
births_ctry <- readRDS("data/processed/births_ctry.rds")
births_rgn <- readRDS("data/processed/births_rgn.rds")
births_lad <- readRDS("data/processed/births_lad.rds") %>%
  aggregate_to_combined_las() # make combined LAs in annual data to match aggregated monthly data


# yearly aggregations of monthly births
births_monthly_ctry <- readRDS("data/processed/births_ctry_yearly_by_month.rds") 
births_monthly_rgn <- readRDS("data/processed/births_rgn_yearly_by_month.rds")
births_monthly_lad <- readRDS("data/processed/births_lad_yearly_by_month.rds")

```


```{r lads}
lad <- births_lad %>%
  left_join(rename(births_monthly_lad, monthly_value = value), by = join_by(gss_code, gss_name, year_ending_date, measure, geography, sex))  %>%
  filter(!is.na(monthly_value)) %>%
  select(-source.x, -source.y, -source_url.x, -source_url.y) %>%
  mutate(diff = monthly_value - value, 
         percdiff = 100*diff/value,
         absdiff = abs(diff),
         abspercdiff = abs(percdiff))
hist(lad$diff)
hist(lad$percdiff)

head(count(lad, abspercdiff) %>% arrange(-abspercdiff), 20)

```

There are some very large differences, but only 16 of the rows have differences above 1%. On further investigation of these 16 rows it looks like a set of births has been recorded in South Somerset in the annual births, but Somerset West and Taunton for monthly ones.


```{r somerset}
filter(lad, abspercdiff > 1) %>% 
  arrange(year_ending_date, gss_code) %>% 
  select(-geography, -sex, -measure) %>%
  print()

```
If we remove South Somerset and Somerset West and Taunton from the dataset the histogram of differences looks more reasonable

```{r no_somerset}
no_somerset <- filter(lad, !gss_code %in% c("E07000189", "E07000246"))
hist(no_somerset$percdiff)

```
Do a check on regions:

```{r region}
rgn <- births_rgn %>%
  left_join(rename(births_monthly_rgn, monthly_value = value), by = join_by(gss_code, gss_name, year_ending_date, sex))  %>%
  filter(!is.na(monthly_value)) %>%
  select(-source.x, -source.y, -source_url.x, -source_url.y, -geography.x, -geography.y) %>%
  mutate(diff = monthly_value - value, 
         percdiff = 100*diff/value,
         absdiff = abs(diff),
         abspercdiff = abs(percdiff))
hist(rgn$diff)
hist(rgn$percdiff)

```
And on country

```{r country}
ctry <- births_ctry %>%
  left_join(rename(births_monthly_ctry, monthly_value = value), by = join_by(gss_code, gss_name, year_ending_date, sex))  %>%
  filter(!is.na(monthly_value)) %>%
  select(-source.x, -source.y, -source_url.x, -source_url.y, -geography.x, -geography.y) %>%
  mutate(diff = monthly_value - value, 
         percdiff = 100*diff/value,
         absdiff = abs(diff),
         abspercdiff = abs(percdiff))
hist(ctry$diff)
hist(ctry$percdiff)


```



